WORCS: A Workflow for Open Reproducible Code in Science

### 2020-11-02 paper/Manuscript_DS.Rmd
R1 C4 fix reference list

---
author:
  - name: Caspar J. van Lissa
    affiliation: 'Utrecht University'
    address: Padualaan 14, 3584CH Utrecht, The Netherlands
    email: c.j.vanlissa@uu.nl
  - name: Andreas M. Brandmaier
    affiliation: 'Max Planck Institute'
  - name: Loek Brinkman
  - name: Anna-Lena Lamprecht
    affiliation: 'Utrecht University'
    email: a.l.lamprecht@uu.nl
  - name: Aaron Peikert
    affiliation: Max Planck Institute
  - name: Marijn E. Struiksma
    affiliation: 'Utrecht University'
  - name: Barbara Vreede
    affiliation: 'Utrecht University'
title: "WORCS: A Workflow for Open Reproducible Code in Science"
short:      "WORKFLOW FOR OPEN SCIENCE"
shorttitle:      "WORKFLOW FOR OPEN SCIENCE"
abstract: |
  Adopting open science
  principles can be challenging and time-intensive, because doing so requires
  substantial conceptual education and training in the use of new tools.
  This paper introduces the Workflow for
  Open Reproducible Code in Science (WORCS): A step-by-step
  procedure that researchers can follow to make a research project open and
  reproducible, accompanied by an open source software implementation that 
  automates these steps.
  The 
  purpose of the workflow is to lower the threshold for 
  adoption of open science principles. 
  It is based on best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies.
  This workflow is primarily targeted towards
  scholars working in R on a private computer,
  conducting research that involves academic prose, analysis code,
  and (optionally) tabular data - but it is flexible enough to accommodate many
  other scenarios.
  To facilitate widespread adoption, the
  WORCS principles are implemented in the R package `worcs`, which offers an
  RStudio 
  project template that is, effectively, a "one-click solution" for R users,
  and utility functions for specific workflow steps.
  This paper addresses the design of the workflow,
  discusses how it meets different 
  standards for open science, and provides examples of the implementation of
  `worcs` in `R`. The source code for the `R` package and manuscript,
  and a list of user examples of WORCS projects, are available at 
  https://github.com/cjvanlissa/worcs.
authornote: "blabla"
keywords:
  # at least one keyword must be supplied
  formatted: [open science, reproducibility, r, dynamic document generation, version control]
  plain:     [open science, reproducibility, r, dynamic document generation, version control]
preamble: >
  \usepackage{amsmath}
bibliography      : ["worcs.bib"]  
urlcolor          : blue
output: rmarkdown::html_document
knit: worcs::cite_all
---

@@

Many researchers are motivated to adopt current best practices for open
science and enjoy these benefits. And yet, the question of how to 
get started can be daunting. Making the transition
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by presenting a simple workflow 
that meets most requirements open science, and is based on best practices:
The *Workflow for Open Reproducible Code in Science* (WORCS).
WORCS is a lightweight approach to open science and
computational reproducibility for scholars working in `R` on a private computer,
conducting research that involves academic prose, analysis code,
and (optionally) tabular data - but it is flexible enough to accommodate many
other scenarios. The workflow (Figure 1) 
constitutes a step-by-step procedure that researchers can follow to make a 
research project open and reproducible.
WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them.
It can also be used to produce work in accordance with best practices
in the absence of top-down support or guidelines.
Although the principles underlying this workflow are universal, WORCS has been
implemented in `R` [@rcoreteamLanguageEnvironmentStatistical2020].
The `worcs` package [@worcspackage2020] offers a project template for RStudio
[@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
a "one click solution". This paper discusses the principles underlying WORCS,
introduces the workflow, addresses how WORCS meets best practices for open science,
and illustrates the use of the `worcs` package.

@@

# Defining Open Science practices

Although Open Science is advocated by many, it does not have a unitary definition
[but see @tennantValuePropositionOpen2020].
One of the most influential operationalisations
of Open Science principles is known as the "TOP-guidelines" [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis; and 8) replication of published results.
WORCS defines the goals of open science in terms of these guidelines, and is
designed to facilitate meeting each of these guidelines, with one exception: We
do not address replication of published results, because replication relates to
the subject of a research project, not to its organization or execution.
We group the remaining seven guidelines into three categories: citation (1),
sharing (2-5), and preregistration (6-7).

@@

WORCS encourages and simplifies the adoption of Open Science principles in daily scientific work. It helps researchers make all research output created throughout the scientific process - not just manuscripts, but data, code, and methods - open and publicly assessible. This enables other researchers to reproduce results, and facilitates cumulative science by allowing others to make direct use of these research objects.

>>>>

---
author:
  name: |
    [Caspar J. Van Lissa](c.j.vanlissa@uu.nl)$^{a,b}$
    Andreas M. Brandmaier$^{c,d}$
    Loek Brinkman$^{b,e}$
    Anna-Lena Lamprecht$^{b,f}$
    Aaron Peikert$^{c,g}$
    Marijn E. Struiksma$^{b,h}$
    Barbara Vreede$^{i}$

  affiliation: | 
      a: Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics  
      b: Open Science Community Utrecht  
      c: Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany  
      d: Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK  
      e: UMC Utrecht  
      f: Utrecht University, Department of Information and Computing Sciences  
      g: Humboldt-Universitaet zu Berlin  
      h: Utrecht University faculty of Humanities, department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS  
      i: University Library Utrecht
title: "WORCS: A Workflow for Open Reproducible Code in Science"
short:      "WORKFLOW FOR OPEN SCIENCE"
shorttitle:      "WORKFLOW FOR OPEN SCIENCE"
abstract: |
  Adopting open science
  principles can be challenging and time-intensive, because doing so requires
  substantial conceptual education and training in the use of new tools.
  This paper introduces the Workflow for
  Open Reproducible Code in Science (WORCS): A step-by-step
  procedure that researchers can follow to make a research project open and
  reproducible, accompanied by an open source software implementation that 
  automates these steps.
  The 
  purpose of the workflow is to lower the threshold for 
  adoption of open science principles. 
  It is based on best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies.
  This workflow is primarily targeted towards
  scholars working in R on a private computer,
  conducting research that involves academic prose, analysis code,
  and (optionally) tabular data - but it is flexible enough to accommodate many
  other scenarios.
  To facilitate widespread adoption, the
  WORCS principles are implemented in the R package `worcs`, which offers an
  RStudio 
  project template that is, effectively, a "one-click solution" for R users,
  and utility functions for specific workflow steps.
  This paper addresses the design of the workflow,
  discusses how it meets different 
  standards for open science, and provides examples of the implementation of
  `worcs` in `R`. The source code for the `R` package and manuscript,
  and a list of user examples of WORCS projects, are available at 
  https://github.com/cjvanlissa/worcs.
authornote: "blabla"
keywords:
  # at least one keyword must be supplied
  formatted: [open science, reproducibility, r, dynamic document generation, version control]
  plain:     [open science, reproducibility, r, dynamic document generation, version control]
preamble: >
  \usepackage{amsmath}
bibliography      : ["worcs.bib"]  
urlcolor          : blue
output: rmarkdown::html_document
---
<!--
, "worcs_packages.bib"
output: 
  rmarkdown::html_document:
    citation_package: biblatex Note: Citations are always handled by pandoc when format is not PDF
knit: worcs::cite_all
csl: apa.csl
csl: apa.csl
-->

@@

Many researchers are motivated to adopt current best practices for open
science and enjoy these benefits. And yet, the question of how to 
get started can be daunting. Making the transition
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by presenting a simple workflow 
that meets most requirements open science, and is based on best practices:
The *Workflow for Open Reproducible Code in Science* (WORCS).
WORCS is a lightweight approach to open science and
computational reproducibility for scholars working in `R` on a private computer,
conducting research that involves academic prose, analysis code,
and (optionally) tabular data - but it is flexible enough to accommodate many
other scenarios. The workflow (Figure 1) 
constitutes a step-by-step procedure that researchers can follow to make a 
research project open and reproducible.
WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them.
It can also be used to produce work in accordance with best practices
in the absence of top-down support or guidelines.
Although the principles underlying this workflow are universal, WORCS has been
implemented in `R` [@rcoreteamLanguageEnvironmentStatistical2020].
The `worcs` package [@worcs] offers a project template for RStudio
[@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
a "one click solution". This paper discusses the principles underlying WORCS,
introduces the workflow, addresses how WORCS meets best practices for open science,
and illustrates the use of the `worcs` package.

@@

# Defining open science practices

Although open science is advocated by many,
it does not have a unitary definition.
Throughout this paper, we adhere to the definition of open science as
"the practice of science in such a way that others can collaborate and contribute,
where research data, lab notes and other research processes are freely available,
under terms that enable reuse, redistribution and reproduction of the research and its underlying data and methods" [@vicente-saezOpenScienceNow2018].
In this definition, the term "science" refers to all scholarly disciplines.
We further define the term *reproducible*, as used in the WORCS acronym, as being able to re-perform "the same analysis with the same code using a different analyst" [@patilVisualToolDefining2019].

The "TOP-guidelines" are one of the most influential concrete operationalisations 
of general open science principles [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science,
the first seven of which "account for reproducibility of the reported results
based on the originating data, and for sharing sufficient information to conduct an independent 
replication" [@nosekPromotingOpenResearch2015a].
In this context, 
reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
These guidelines are: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis.
The eighth criterion is "not formally a transparency standard for authors", but "addresses 
journal guidelines for consideration of independent replications for publication".
WORCS defines the goals of open science in terms of the first seven TOP-guidelines,
and the workflow and its `R` implementation are designed to facilitate compliance therewith.
We group these guidelines into three categories: citation (1),
sharing (2-5), and preregistration (6-7).

@@

WORCS encourages and simplifies the adoption of open science principles in daily scientific work. It helps researchers make all research output created throughout the scientific process - not just manuscripts, but data, code, and methods - open and publicly assessible. This enables other researchers to reproduce results, and facilitates cumulative science by allowing others to make direct use of these research objects.

### 2020-11-02 paper/reviews_ds.Rmd
R1 C4 fix reference list

\RC{There is no user evaluation, and therefore all the claims about worcs are unsupported by evidence. If the authors had included a compelling evaluation I would be in favor of accepting the paper.}

@@

\RC{This paper describes worcs, a package and workflow to help make R code reproducible according to the TOP guidelines.

@@

\RC{Unsupported claims (beyond the ones derived from an absence of evaluation): The paper claims in many parts that worcs helps make an entire research project Open and Reproducible. However, at the same time there is also a claim for not supporting replication. Since replication is generally understood as a necessary steps towards reproducibility, I find this kind of contradictory. Maybe the claims can be modified stating that worcs helps documentation and understanding, hence supporting FAIR.}

Response: We appreciate the Reviewer's comment, which seems to be rooted in a different definition of the terms "replication" and "reproducibility" than was intended by us. The lack of consistent definitions for these terms has been lamented by other authors (e.g., Patil, Peng, & Leek, 2019; Plesser, 2018). As WORCS aims to meet the requirements set out in the TOP-guidelines, we adhere to the definitions of reproducibility and replication as used in the TOP-guidelines. This definition is formalized by Patil and colleagues: "We define reproducibility as re-performing the same analysis with the same code using a different analyst, and we define replicability as re-performing the experiment and collecting new data". The TOP-guidelines state that the first seven guidelines are intended to ensure the "reproducibility of the reported results based on the originating data". The eighth criterion is "not formally a transparency standard for authors, [but] addresses journal guidelines for consideration of independent replications for publication."

We had taken notice of this definitional quagmire, but decided not to dwell on it in the present paper. This reviewer comment has made it abundantly clear that we need to explicitly define our terms. Thus, in order to address this comment, the revised manuscript now reads: REF.

We wish to fully address this comment, including the criticism of unsupported claims - but it is a bit difficult to do so, given the fundamental misunderstanding about what constitutes "reproducibility". We have stated that worcs helps make an entire research project open and reproducible. It does so by:
1) Providing a conceptual workflow that addresses all steps required to make a project open and reproducible
2) Addressing common bottlenecks, such as the inability to share primary data, and offering solutions to overcome these hurdles
3) Version controlling a project on Git(Hub)
4) Creating a fully reproducible RMarkdown document
5) Managing the package dependencies

The Reviewer, by contrast, proposes that the main contribution of worcs is to help documentation and understanding. It is not immediately clear to us which components of worcs address these two points - except perhaps the readme.md that is created upon initialization of a new project.
\Assignment{CJ, ALL}
\WorkInProgress
\Hard

@@

Response: Citation should be formatted as software.

>>>>

\RC{C1 There is no user evaluation, and therefore all the claims about worcs are unsupported by evidence. If the authors had included a compelling evaluation I would be in favor of accepting the paper.}

@@

\RC{C2 This paper describes worcs, a package and workflow to help make R code reproducible according to the TOP guidelines.

@@

\RC{C3 Unsupported claims (beyond the ones derived from an absence of evaluation): The paper claims in many parts that worcs helps make an entire research project Open and Reproducible. However, at the same time there is also a claim for not supporting replication. Since replication is generally understood as a necessary steps towards reproducibility, I find this kind of contradictory. Maybe the claims can be modified stating that worcs helps documentation and understanding, hence supporting FAIR.}

Response: We understand that our writing comes across as contradictory,
and this apparent contradiction arises from a lack of explicit definitions of the terms "replication" and "reproducibility" in the original manuscript.
We are aware of the oft-lamented lack of consistent definitions for these terms (e.g., Patil, Peng, & Leek, 2019; Plesser, 2018).
In the Revision, we have attempted to address this lack of clarity by including explicit definitions for these terms:

> Throughout this paper, we adhere to the definition of open science as
"the practice of science in such a way that others can collaborate and contribute,
where research data, lab notes and other research processes are freely available,
under terms that enable reuse, redistribution and reproduction of the research and its underlying data and methods" [@vicente-saezOpenScienceNow2018].  
[...] reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].

After making these definitions explicit, it becomes clear that the Reviewer's definition of replication as "a necessary steps towards reproducibility" is not consistent with the present paper.

We further wish to clarify that the fact that we address guidelines 1-7, but not guideline 8 ("replication") is consistent with the intended purpose of the TOP-guidelines.
Specifically, the first seven guidelines are intended to ensure the "reproducibility of the reported results based on the originating data" [direct quote from @nosekPromotingOpenResearch2015a]. The eighth criterion, by contrast, is "not formally a transparency standard for authors, [but] addresses journal guidelines for consideration of independent replications for publication" (again a direct quote).
To avoid further confusion or controversy, we now address this distinction between the first 7 guidelines and the final 8th explicitly in the revised manuscript:

> These guidelines describe eight standards for open science,
the first seven of which "account for reproducibility of the reported results
based on the originating data, and for sharing sufficient information to conduct an independent 
replication" [@nosekPromotingOpenResearch2015a].
In this context, 
reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
These guidelines are: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis.
The eighth criterion is "not formally a transparency standard for authors", but "addresses 
journal guidelines for consideration of independent replications for publication".
WORCS defines the goals of open science in terms of the first seven TOP-guidelines,
and the workflow and its `R` implementation are designed to facilitate compliance therewith.

<!--We wish to fully address this comment, including the criticism of unsupported claims - but it is a bit difficult to do so, given the fundamental misunderstanding about what constitutes "reproducibility". We have stated that worcs helps make an entire research project open and reproducible. It does so by:
1) Providing a conceptual workflow that addresses all steps required to make a project open and reproducible
2) Addressing common bottlenecks, such as the inability to share primary data, and offering solutions to overcome these hurdles
3) Version controlling a project on Git(Hub)
4) Creating a fully reproducible RMarkdown document
5) Managing the package dependencies

The Reviewer, by contrast, proposes that the main contribution of worcs is to help documentation and understanding. It is not immediately clear to us which components of worcs address these two points - except perhaps the readme.md that is created upon initialization of a new project.-->

\Assignment{CJ, ALL}
\Done
\Hard

@@

Response: We apologize; it appears that there were numerous errors in the parsing of the reference list after rendering the manuscript to HTML - the preferred format of Data Science. The reference is indeed to the R-package. In the revised manuscript, we have repaired the references so that it is clear that this is a software package, et cetera.
